### 2014-11-02 paper/experiments.tex
fixing some grammar and spelling issues

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. Our solution and Yahoo S4\cite{neumeyer2010s4} EventProducer reads a file containing over 7500000 ecg records and push events with multiple threads. Twitter Storm\cite{twitterStorm} does not allow user threads to push data. So we used the Spout thread to send messages into the system. EventReceiver  receives the ECG events, process them and calculate heart rate interval periodically.  
\subsubsection{Throughput}
We execute our system with 1, 2 and 4  EventReceivers for each system and measured the throughput, load average and network bandwidth  for each case. Throughput was measured at the EventProducer by calculating the total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

@@

As shown in the Figure \ref{throuput} our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}. However by looking the Figure \ref{throuput} one might think even our solution does not scalable although it perform high. In order to examine the reason behind this we looked into the network bandwidth and CPU load average values.

@@

By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers. In only one receiver case, high CPU load average at receiver indicates it has utilized available CPU power. Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth. Therefore it can not increase the throughput without increasing network bandwidth although more CPU power available at receiver nodes. Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver since load is shared among the receivers. However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage. If a system does not make use of the available resources then adding more resources won't scale up the system. As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient 
message parse technique.

\subsubsection{Efficiency of Message Serialization}
After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from EventProducer to EventReceiver. Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string. So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID). Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message. By reducing 20 information bytes we can calculated the overhead bytes for each system. Figure \ref{efficiency} compares these numbers.

@@

This application predicts load averages using the previous values and a given machine learning algorithm. We implement this logic using nodes as shown in the Figure \ref{multigraph}. First producer reads the data file and send events to the avgCal node which calculate the last minute average and send the same event to both plugPredict and housePredict processors. Both plugPredict and housePredict processors predict the next values and send events to receivers. The original problem only requires to send those prediction events in 30s intervals. But we sent a prediction event for each messages to observe how the system works with high load.

As in the earlier case we conduct our experiments using one producer and multiplying the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer. We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively. We measured the throughput,  network bandwidth and the CPU load average at the producer to examine the scalability of the system. Since the one receiver unit throughput greater than that of the storm one node scenario we only used our system for this experiment.  Results are shown in the Figure \ref{scalability}.

@@

For one receiver unit case, we observed very high load average and network bandwidth usage at avgCal node since it sends messages to two nodes. Then as shown in the figure \ref{scalability} we could linearly scale up the system by adding more receiver units to add more cpu power to system. When increasing the receiver units, we can observe throughput of the system increases proportional to number of receiving units. CPU load average of 13 at the producer indicates it has utilized all available CPU power and we need to add more producers to scale this system further up.

>>>>

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. Our solution and Yahoo S4\cite{neumeyer2010s4} \textit{EventProducer} reads a file containing over 7500000 ecg records and pushes events with multiple threads. Twitter Storm\cite{twitterStorm} does not allow user threads to push data. So we used the \textit{spout} thread to send messages into the system. \textit{EventReceiver}  receives the ECG events, process them and calculate heart rate interval periodically.  
\subsubsection{Throughput}
We executed our system with 1, 2 and 4  \textit{EventReceivers} for each system and measured the throughput, load average and network bandwidth  for each case. Throughput was measured at the \textit{EventProducer} by calculating the total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

@@

As shown in the Figure \ref{throuput} our solution outperforms Twitter storm\cite{twitterStorm} and Yahoo S4\cite{neumeyer2010s4}. However by looking the Figure \ref{throuput} one might think even our solution is not scalable although it performs high. In order to examine the reason behind this we looked into the network bandwidth and CPU load average values.

@@

By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers. In only one receiver case, high CPU load average at receiver indicates it has utilized available CPU power. Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth. Therefore it can not increase the throughput without increasing network bandwidth although more CPU power is available at receiver nodes. Having receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver, since load is shared among the receivers. However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage. If a system does not make use of the available resources then adding more resources won't scale up the system. As explained in earlier, our solution utilities all available resources by using parallel tcp connections and having an efficient 
message parse technique.

\subsubsection{Efficiency of Message Serialization}
After measuring the throughput we examined the  total amount of extra bytes each system sends to transfer the information from \textit{EventProducer} to \textit{EventReceiver}. Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string. So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID). Then by using the throughput and the bandwidth usage of the system we calculated the total number of bytes each system uses at network layer to send this message. By reducing 20 information bytes we calculated the overhead bytes for each system. Figure \ref{efficiency} compares these numbers.

@@

This application predicts load averages using previous values and a machine learning algorithm. We implemented this logic using nodes as shown in the Figure \ref{multigraph}. First producer reads the data file and send events to the \textit{avgCal} node which calculates the last minute average and sends the same event to both \textit{plugPredict} and \textit{housePredict} processors. Both \textit{plugPredict} and \textit{housePredict} processors predict the next values and send events to receivers. The original problem only requires to send those prediction events in 30s intervals. But we used a prediction event for each messages to observe how the system works with high load.

As in the earlier case, we conducted our experiments using one producer and multiplying the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer. We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively. We measured the throughput,  network bandwidth and the CPU load average at the producer to examine the scalability of the system. Since the one receiver unit throughput greater than that of the storm one node scenario, we only used our system for this experiment.  Results are shown in the Figure \ref{scalability}.

@@

For one receiver unit case, we observed very high load average and network bandwidth usage at \textit{avgCal} node since it sends messages to two nodes. Then as shown in the figure \ref{scalability} we could linearly scale up the system by adding more receiver units to add more cpu power to the system. When increasing the receiver units, we could observe throughput of the system increases proportional to number of receiving units. CPU load average of 13 at the producer indicates it has utilized all available CPU power and it is required to add more producers to scale this system further up.

### 2014-11-02 paper/system_architecture.tex
fixing some grammar and spelling issues

\section{System Design}
In this section we elaborate the underlying system design which is capable of processing 2.4 million messages per second. We emphasize three features of this design. Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages. All processes share the same set of connections which we called as the bridge. The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we show later our experiments shows this approach perform several times better than the other systems despite above overhead. Secondly we have implemented OutputStream and InputStream interfaces on top of java.nio API. Using these input and output streams we provide DataOutput and DataInput APIs to directly serialize message data to binary format without sending any metadata along the message. Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time. When a process sends the first message all these objects are created and it is a one time cost. Rest of this section further explain design techniques we used to realise these concepts.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involves in sending a message from one process to another. Lets assume there is a message at a client side process and that process wants to send this message to another process. First client process sends this message to its ElementContainer. ElementContainer holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream decides which node to send this message and passes the message along the target node details to ConnectionManager which holds ClientConnections for each node. ConnectionManager picks the correct ClientConnection for target node and sends the message to ClientConnection. ClientConnection picks an available DataOutput from its pool of DataOutputs and invokes the serialize method of the message to send that to server side.

At the server side there is a set of ServerTasks which read receiving messages from pool of DataInput streams available in the ServerConnection (we register these connections at the connections creating time). Once a server task receives a binary message, it creates the event using the sending process id to identify the event type. Then it pass this message to the WorkerContainer which holds all processes. Finally WorkerContainer dispatch this message to correct process using receiving process id. Message ordering happens at the process level if required. This illustrates how we have shared a pool of tcp connections among process and next we show how we have implemented OutputStream and InputStream with non blocking I/O API without  creating objects dynamically.
\begin{figure}[!tii]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented OutputStream interface using our DataWriter. DataWriter contains a ByteBuffer. When a higher layer method writes data to DataOuput, DataWriter put those bytes to its byte buffer. When a selection event occurs, Datawriter writes its byte buffer to underline SocketChannel. This way we use one byte buffer as an intermediary to create an output stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented InputStream interface using our DataReader. Similar to client side, DataReader has a ByteBuffer. When a selection event occurs DataReader input bytes from the SocketChannel to its byte buffer. When a ServerTask reads data through DataInput, it reads data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}

>>>>

\section{System Design}
In this section we elaborate the underlying system design which is capable of processing 2.4 million messages per second. We emphasize three features of this design. Firstly between any two nodes there is a set of tcp connections (typically 20 connections)  to send messages. All processes share the same set of connections which we called as the bridge. The advantage of this approach compared to having tcp connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we need to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we show later our experiments shows this approach perform several times better than the other systems despite above overhead. Secondly we have implemented \textit{OutputStream} and \textit{InputStream} interfaces on top of java non blocking I/O API. Using these input and output streams we provide \textit{DataOutput} and \textit{DataInput} APIs to directly 
serialize message data to binary format without sending any metadata along the message. Finally except messages itself all the other objects used to serialize the messages and parse the messages are not created at message processing time. When a process sends the first message all these objects are created and it is a one time cost. Rest of this section further explain design techniques we used to realise these concepts.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involves in sending a message from one process to another. Lets assume there is a message at a client side process and that process wants to send this message to another process. First client process sends this message to its \textit{ElementContainer}. \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream decides which node to send this message and passes the message along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node. \textit{ConnectionManager} sends message to correct \textit{ClientConnection} using target node. \textit{ClientConnection} picks an available \textit{DataOutput} from its pool of \textit{DataOutputs} and invokes the serialize method of the message to send that to server side.

At the server side there is a set of \textit{ServerTask}s which read receiving messages from pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time). Once a \textit{ServerTask} receives a binary message, it creates the event using the sending process id to identify the event type. Then it pass this message to the \textit{WorkerContainer} which holds all processes. Finally \textit{WorkerContainer} dispatch this message to correct process using receiving process id. Message ordering happens at the process level if required. This illustrates how we have shared a pool of tcp connections among process and next we show how we have implemented \textit{OutputStream} and \textit{InputStream} with non blocking I/O API without  creating objects dynamically.
\begin{figure}[!tii]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented \textit{OutputStream} interface using our \textit{DataWriter}. \textit{DataWriter} contains a \textit{ByteBuffer}. When a higher layer method writes data to \textit{DataOuput}, \textit{DataWriter} puts those bytes to its byte buffer. When a selection event occurs, \textit{Datawriter} writes its byte buffer to underline \textit{SocketChannel}. This way only one byte buffer is used as an intermediary to create an output stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented \textit{InputStream} interface using our \textit{DataReader}. Similar to client side, \textit{DataReader} has a \textit{ByteBuffer}. When a selection event occurs \textit{DataReader} inputs bytes from the \textit{SocketChannel} to its byte buffer. When a \textit{ServerTask} reads data through \textit{DataInput}, it reads data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}
